---
title: "ASH-based wavelet shrinkage for heteroskedastic models"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{smashr demo}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, echo = FALSE, message = FALSE}
knitr::opts_chunk$set(collapse = TRUE,comment = "#",fig.align = "center",
                      fig.cap = "&nbsp;",dpi = 120,results = "hold")
```

In this vignette, we assume a simple parametric model of the form

$$
y_i = \mu_i + \epsilon_i,
$$

where $\epsilon_i \sim N(0,\sigma_i^2)$, and both $\mu$ and $\sigma$
are unknowns. We use the "smash" procedure (SMoothing via Adaptive
SHrinkage) to estimate both the mean and the variances. Here we
present a brief demonstration of the method.

We first look at mean estimation, which is our primary focus. A sample
mean function is presented, as well as a couple of different variance
functions. Our method is compared against a few other simple methods.

## Set up environment

We begin by loading the smashr, EbayesThresh and wavethresh packages.

```{r load-pkgs, message=FALSE}
library(smashr)
library(EbayesThresh)
library(wavethresh)
```

## Preparations

First, we define the mean function used to simulate the data.

```{r define-spike-function}
spike.f <- function(x) (0.75 * exp(-500 * (x - 0.23)^2) +
  1.5  * exp(-2000  * (x - 0.33)^2) +
  3    * exp(-8000  * (x - 0.47)^2) +
  2.25 * exp(-16000 * (x - 0.69)^2) +
  0.5  * exp(-32000 * (x - 0.83)^2))
n    <- 1024
t    <- 1:n/n
mu.s <- spike.f(t)
```

Define a few other functions which will be used in the code chunks
below.

```{r define-functions}
mse    <- function(x, y) mean((x - y)^2)
l2norm <- function(x) sum(x^2)
mise   <- function(x, y, r)
  10000 * mean(apply(x - rep(1, r) %o% y, 1, l2norm)/l2norm(y))
sig.est.func <- function(x, n)
  sqrt(2/(3 * (n - 2)) * sum((1/2 * x[1:(n - 2)] - x[2:(n - 1)] +
  1/2 * x[3:n])^2))
```

Define a function for the default wavelet thresholding method.

```{r define-wavelet-default}
waveti.u <- function(x, filter.number = 10, family = "DaubLeAsymm",
                     min.level = 3, noise.level) {
    TT = length(x)
    thresh = noise.level * sqrt(2 * log(TT))
    x.w = wavethresh::wd(x, filter.number, family, type = "station")
    x.w.t = threshold(x.w, levels = (min.level):(x.w$nlevels - 1),
	                  policy = "manual", value = thresh, type = "hard")
    x.w.t.r = AvBasis(convert(x.w.t))
    return(x.w.t.r)
}
```

Define another function for the default EbayesThresh method.

```{r define-ebayesthresh-default}
waveti.ebayes <- function(x, filter.number = 10, family = "DaubLeAsymm",
                          min.level = 3, noise.level) {
    n = length(x)
    J = log2(n)
    x.w = wd(x, filter.number, family, type = "station")
    for (j in min.level:(J - 1)) {
        x.pm = ebayesthresh(accessD(x.w, j), sdev = noise.level)
        x.w = putD(x.w, j, x.pm)
    }
    mu.est = AvBasis(convert(x.w))
    return(mu.est)
}
```

For the first demonstration, define the mean and variance functions,
and set the signal to noise ratio.

```{r mean-and-variance-functions}
mu.t <- (1 + mu.s)/5
rsnr <- sqrt(1)
var1 <- rep(1, n)
var2 <- (1e-04 + 4 * (exp(-550 * (t - 0.2)^2) + exp(-200 * (t - 0.5)^2) +
          exp(-950 * (t - 0.8)^2)))/1.35 
```

## Constant variance example

We first look at the case of constant variance.

```{r constant-variance-example}
set.seed(327)
sigma.ini <- sqrt(var1)
sigma.t   <- sigma.ini/mean(sigma.ini) * sd(mu.t)/rsnr^2
X.s       <- matrix(rnorm(10 * n, mu.t, sigma.t), nrow = 10, byrow = TRUE)
mu.est    <- apply(X.s, 1, smash.gaus)
mu.est.tivar.ash <- apply(X.s, 1, ti.thresh, method = "smash")
mu.est.tivar.mad <- apply(X.s, 1, ti.thresh, method = "rmad")
mu.est.ti        <- matrix(0, 10, n)
mu.est.ti.ebayes <- matrix(0, 10, n)
for (i in 1:10) {
  sig.est = sig.est.func(X.s[i, ], n)
  mu.est.ti[i, ] = waveti.u(X.s[i, ], noise.level = sig.est)
  mu.est.ti.ebayes[i, ] = waveti.ebayes(X.s[i, ], noise.level = sig.est)
}
```

Assess the accuracy of the results.

```{r calc-mise}
cat("SMASH:",mise(t(mu.est), mu.t, 10),"\n")
cat("TI thresholding with variance estimated from smash:",
    mise(t(mu.est.tivar.ash), mu.t, 10),"\n")
cat("TI thresholding with variance estimated from running MAD:",
    mise(t(mu.est.tivar.mad), mu.t, 10),"\n")
cat("TI thresholding with constant variance (estimated):",
	mise(mu.est.ti, mu.t, 10),"\n")
cat("EBayes with constant variance (estimated):",
	mise(mu.est.ti.ebayes, mu.t, 10),"\n")
```

Plot the estimated mean functions against the ground-truth function
(in black).

```{r plots-1, fig.width=5, fig.height=4}
plot(mu.t, xlab = "", ylab = "", type = "l")
lines(mu.est[, 1], col = 2)
lines(mu.est.tivar.mad[, 1], col = 3)
lines(mu.est.ti[1, ], col = 4)
lines(mu.est.ti.ebayes[1, ], col = 6)
legend("topright", legend = c("smash", "ti_rmad", "ti_homo", "ebayes_homo"),
       fill = c(2, 3, 4, 6))
```

## Non-constant variance example

Generate the data for this example.

```{r non-constant-variance-example}
sigma.ini = sqrt(var2)
sigma.t = sigma.ini/mean(sigma.ini) * sd(mu.t)/rsnr^2
set.seed(327)
X.s = matrix(rnorm(10 * n, mu.t, sigma.t), nrow = 10, byrow = TRUE)
mu.est = apply(X.s, 1, smash.gaus)
mu.est.tivar.ash = apply(X.s, 1, ti.thresh, method = "smash")
mu.est.tivar.mad = apply(X.s, 1, ti.thresh, method = "rmad")
mu.est.ti = matrix(0, 10, n)
mu.est.ti.ebayes = matrix(0, 10, n)
for (i in 1:10) {
    sig.est = sig.est.func(X.s[i, ], n)
    mu.est.ti[i, ] = waveti.u(X.s[i, ], noise.level = sig.est)
    mu.est.ti.ebayes[i, ] = waveti.ebayes(X.s[i, ], noise.level = sig.est)
}
```

Assess the accuracy of the results.

```{r calc-mise-2}
cat("SMASH:",mise(t(mu.est), mu.t, 10),"\n")
cat("TI thresholding with variance estimated from smash:",
    mise(t(mu.est.tivar.ash), mu.t, 10),"\n")
cat("TI thresholding with variance estimated from running MAD:",
    mise(t(mu.est.tivar.mad), mu.t, 10),"\n")
cat("TI thresholding with constant variance (estimated):",
	mise(mu.est.ti, mu.t, 10),"\n")
cat("EBayes with constant variance (estimated):",
	mise(mu.est.ti.ebayes, mu.t, 10) ,"\n")
```

Plot the estimated mean functions against the ground-truth function
(in black).

```{r plots-2, fig.width=5, fig.height=4}
plot(mu.t, xlab = "", ylab = "", type = "l")
lines(mu.est[, 1], col = 2)
lines(mu.est.tivar.mad[, 1], col = 3)
lines(mu.est.ti[1, ], col = 4)
lines(mu.est.ti.ebayes[1, ], col = 6)
legend("topright", legend = c("smash", "ti_rmad", "ti_homo", "ebayes_homo"),
       fill = c(2, 3, 4, 6))
```

## Estimating the variance function

Next we look at variance estimation. Because variance estimation is
much harder, we use a larger sample size. In this example, we set $n =
4096$.

```{r estimated-variance-example}
n = 2^12
t = 1:n/n
mu.s = spike.f(t)
```

Define the test functions and generate the data.

```{r est-var-functions}
pos = c(0.1, 0.13, 0.15, 0.23, 0.25, 0.4, 0.44, 0.65, 0.76, 0.78, 0.81)
hgt = c(4, 5, 3, 4, 5, 4.2, 2.1, 4.3, 3.1, 5.1, 4.2)
wth = c(0.005, 0.005, 0.006, 0.01, 0.01, 0.03, 0.01, 0.01, 0.005, 0.008, 0.005)
mu.b = rep(0, n)
for (j in 1:length(pos)) {
    mu.b = mu.b + hgt[j]/((1 + (abs(t - pos[j])/wth[j]))^4)
}
dop.f = function(x) sqrt(x * (1 - x)) * sin((2 * pi * 1.05)/(x + 0.05))
mu.dop = dop.f(t)

var1.ini = ((3 - 20 * t) * (t >= 0 & t < 0.1) +
  (20 * t - 1) * (t >= 0.1 & t < 0.25) +
  (4 + (1 - 4 * t) * 18/19) * (t >= 0.25 & t < 0.725) + 
  (2.2 + 10 * (t - 0.725)) * (t >= 0.725 & t < 0.89) +
  (3.85 - 85 * (t - 0.89)/11) * (t >= 0.89 & t <= 1))
var1 = var1.ini/sqrt(var(var1.ini))
var2.ini = (1 + 4 * (exp(-550 * (t - 0.2)^2) +
  exp(-200 * (t - 0.5)^2) + exp(-950 * (t - 0.8)^2)))
var2 = var2.ini/sqrt(var(var2.ini))
var3 = mu.b/sqrt(var(mu.b))
var4 = (mu.dop + 2)/sqrt(var(mu.dop))

sigma.t.1 = sqrt(var1)
sigma.t.2 = sqrt(var2)
sigma.t.3 = sqrt(var3)
sigma.t.4 = sqrt(var4)

set.seed(327)
X.s.1 = rnorm(n, 0, sigma.t.1)
set.seed(327)
X.s.2 = rnorm(n, 0, sigma.t.2)
set.seed(327)
X.s.3 = rnorm(n, 0, sigma.t.3)
set.seed(327)
X.s.4 = rnorm(n, 0, sigma.t.4)
```

Fit the SMASH model to these data.

```{r est-var-smash}
var.est.1 = smash.gaus(X.s.1, v.est = TRUE)
var.est.2 = smash.gaus(X.s.2, v.est = TRUE)
var.est.3 = smash.gaus(X.s.3, v.est = TRUE)
var.est.4 = smash.gaus(X.s.4, v.est = TRUE)
```

Evaluate the fit of these models.

```{r est-var-mse}
mse(var.est.1, sigma.t.1^2)
mse(var.est.2, sigma.t.2^2)
mse(var.est.3, sigma.t.3^2)
mse(var.est.4, sigma.t.4^2)
```

These plots show the ground-truth variance (black) against the
variance functions estimated by SMASH (red).

```{r plots-3, fig.width=5.5, fig.height=5}
par(mfrow = c(2, 2))
plot(sigma.t.1^2, xlab = "", ylab = "", type = "l", main = "V1")
lines(var.est.1, col = 2)
plot(sigma.t.2^2, xlab = "", ylab = "", type = "l", main = "V2")
lines(var.est.2, col = 2)
plot(sigma.t.3^2, xlab = "", ylab = "", type = "l", main = "V3")
lines(var.est.3, col = 2)
plot(sigma.t.4^2, xlab = "", ylab = "", type = "l", main = "V4")
lines(var.est.4, col = 2) 
```

## Unevenly spaced data

In this section, we examine how this wavelet-based approach handles
unevenly spaced data. Wavelet methods are not optimized for unevenly
spaced data, but still works. We first look at the case when the data are
normally spaced.

```{r unevenly-spaced}
n = 1024
t = sort(rnorm(1024, 0, 1))
t = (t - min(t))/(max(t) - min(t))
mu.s = spike.f(t)
```

Define mean function, signal-to-noise ratio, and the variance functions.

```{r unevenly-spaced-functions}
mu.t = (1 + mu.s)/5
rsnr = sqrt(1)
var1 = rep(1, n)
var2 = (1e-04 + 4 * (exp(-550 * (t - 0.2)^2) + exp(-200 * (t - 0.5)^2) +
  exp(-950 * (t - 0.8)^2)))/1.35
```

### Unevenly spaced data with constant variance

First, we look at the case of constant variance.

```{r unevenly-spaced-const-var}
sigma.ini = sqrt(var1)
sigma.t = sigma.ini/mean(sigma.ini) * sd(mu.t)/rsnr^2
set.seed(327)
X.s = matrix(rnorm(10 * n, mu.t, sigma.t), nrow = 10, byrow = TRUE)
```

Apply the methods on the data.

```{r unevenly-spaced-fit-models}
mu.est = apply(X.s, 1, smash.gaus)
mu.est.tivar.ash = apply(X.s, 1, ti.thresh, method = "smash")
mu.est.tivar.mad = apply(X.s, 1, ti.thresh, method = "rmad")
mu.est.ti = matrix(0, 10, n)
mu.est.ti.ebayes = matrix(0, 10, n)
for (i in 1:10) {
    sig.est = sig.est.func(X.s[i, ], n)
    mu.est.ti[i, ] = waveti.u(X.s[i, ], noise.level = sig.est)
    mu.est.ti.ebayes[i, ] = waveti.ebayes(X.s[i, ], noise.level = sig.est)
}
```

Assess the accuracy of the estimates.

```{r calc-mise-3}
cat("SMASH:",mise(t(mu.est), mu.t, 10),"\n")
cat("TI thresholding with variance estimated from smash:",
    mise(t(mu.est.tivar.ash), mu.t, 10),"\n")
cat("TI thresholding with variance estimated from running MAD:",
    mise(t(mu.est.tivar.mad), mu.t, 10),"\n")
cat("TI thresholding with constant variance (estimated):",
	mise(mu.est.ti, mu.t, 10),"\n")
cat("EBayes with constant variance (estimated):",
	mise(mu.est.ti.ebayes, mu.t, 10),"\n")
```

Assess the accuracy over a grid.

```{r calc-mise-grid}
xgrid = 1:n/n
mu.est.inter = matrix(0, 10, n)
mu.est.tivar.ash.inter = matrix(0, 10, n)
mu.est.tivar.mad.inter = matrix(0, 10, n)
mu.est.ti.inter = matrix(0, 10, n)
mu.est.ti.ebayes.inter = matrix(0, 10, n)
for (i in 1:10) {
    mu.est.inter[i, ] = approx(t, mu.est[, i], xgrid)$y
    mu.est.tivar.ash.inter[i, ] = approx(t, mu.est.tivar.ash[, i], xgrid)$y
    mu.est.tivar.mad.inter[i, ] = approx(t, mu.est.tivar.mad[, i], xgrid)$y
    mu.est.ti.inter[i, ] = approx(t, mu.est.ti[i, ], xgrid)$y
    mu.est.ti.ebayes.inter[i, ] = approx(t, mu.est.ti.ebayes[i, ], xgrid)$y
    
}
mu.t.inter = (spike.f(xgrid) + 1)/5
mise(mu.est.inter, mu.t.inter, 10)
mise(mu.est.tivar.ash.inter, mu.t.inter, 10)
mise(mu.est.tivar.mad.inter, mu.t.inter, 10)
mise(mu.est.ti.inter, mu.t.inter, 10)
mise(mu.est.ti.ebayes.inter, mu.t.inter, 10) 
```

Plot the estimated mean functions against the ground-truth function
(in black).

```{r plots-4, fig.width=5, fig.height=4}
par(mfrow = c(1, 1))
plot(t, mu.t, type = "l")
lines(t, mu.est[, 1], col = 2)
lines(t, mu.est.tivar.mad[, 1], col = 3)
lines(t, mu.est.ti[1, ], col = 4)
lines(t, mu.est.ti.ebayes[1, ], col = 6)
legend("topright", legend = c("smash", "ti_rmad", "ti_homo", "ebayes_homo"),
       fill = c(2, 3, 4, 6))
```

### Unevenly spaced data with non-constant variance

We then look at the case of non-constant variance:

```{r unevenly-spaced-nonconstant-var}
sigma.ini = sqrt(var2)
sigma.t = sigma.ini/mean(sigma.ini) * sd(mu.t)/rsnr^2
set.seed(327)
X.s = matrix(rnorm(10 * n, mu.t, sigma.t), nrow = 10, byrow = TRUE)
```

```{r unevenly-spaced-nonconstant-var-fit-models}
mu.est = apply(X.s, 1, smash.gaus)
mu.est.tivar.ash = apply(X.s, 1, ti.thresh, method = "smash")
mu.est.tivar.mad = apply(X.s, 1, ti.thresh, method = "rmad")
mu.est.ti = matrix(0, 10, n)
mu.est.ti.ebayes = matrix(0, 10, n)
for (i in 1:10) {
    sig.est = sig.est.func(X.s[i, ], n)
    mu.est.ti[i, ] = waveti.u(X.s[i, ], noise.level = sig.est)
    mu.est.ti.ebayes[i, ] = waveti.ebayes(X.s[i, ], noise.level = sig.est)
}
```

Assess the accuracy of the estimates.

```{r calc-mise-4}
cat("SMASH:",mise(t(mu.est), mu.t, 10),"\n")
cat("TI thresholding with variance estimated from smash:",
    mise(t(mu.est.tivar.ash), mu.t, 10),"\n")
cat("TI thresholding with variance estimated from running MAD:",
    mise(t(mu.est.tivar.mad), mu.t, 10),"\n")
cat("TI thresholding with constant variance (estimated):",
	mise(mu.est.ti, mu.t, 10),"\n")
cat("EBayes with constant variance (estimated):",
	mise(mu.est.ti.ebayes, mu.t, 10),"\n")
```
	
Assess the accuracy over a grid.
	
```{r calc-mise-grid-2}
xgrid = 1:n/n
mu.est.inter = matrix(0, 10, n)
mu.est.tivar.ash.inter = matrix(0, 10, n)
mu.est.tivar.mad.inter = matrix(0, 10, n)
mu.est.ti.inter = matrix(0, 10, n)
mu.est.ti.ebayes.inter = matrix(0, 10, n)
for (i in 1:10) {
    mu.est.inter[i, ] = approx(t, mu.est[, i], xgrid)$y
    mu.est.tivar.ash.inter[i, ] = approx(t, mu.est.tivar.ash[, i], xgrid)$y
    mu.est.tivar.mad.inter[i, ] = approx(t, mu.est.tivar.mad[, i], xgrid)$y
    mu.est.ti.inter[i, ] = approx(t, mu.est.ti[i, ], xgrid)$y
    mu.est.ti.ebayes.inter[i, ] = approx(t, mu.est.ti.ebayes[i, ], xgrid)$y
}
mu.t.inter = (spike.f(xgrid) + 1)/5
mise(mu.est.inter, mu.t.inter, 10)
mise(mu.est.tivar.ash.inter, mu.t.inter, 10)
mise(mu.est.tivar.mad.inter, mu.t.inter, 10)
mise(mu.est.ti.inter, mu.t.inter, 10)
mise(mu.est.ti.ebayes.inter, mu.t.inter, 10) 
```

Plot the estimated mean functions against the ground-truth function
(in black).

```{r plots-5, fig.width=5, fig.height=4}
par(mfrow = c(1, 1))
plot(t, mu.t, xlab = "", ylab = "", type = "l")
lines(t, mu.est[, 1], col = 2)
lines(t, mu.est.tivar.mad[, 1], col = 3)
lines(t, mu.est.ti[1, ], col = 4)
lines(t, mu.est.ti.ebayes[1, ], col = 6)
legend("topright", legend = c("smash", "ti_rmad", "ti_homo", "ebayes_homo"),
       fill = c(2, 3, 4, 6)) 
```

## Poisson-distributed data example

Now we consider the case in which the data are distributed according
to a Poisson process.

```{r poisson-example}
n    <- 1024
t    <- c(0, rexp(n - 1, 1))
t    <- cumsum(t)
t    <- (t - min(t))/(max(t) - min(t))
mu.s <- spike.f(t)
```

Next, define the mean and variance functions.

```{r poisson-example-mean-and-variance}
mu.t <- (1 + mu.s)/5
rsnr <- sqrt(1)
var1 <- rep(1, n)
var2 <- (1e-04 + 4 * (exp(-550 * (t - 0.2)^2) +
        exp(-200 * (t - 0.5)^2) + exp(-950 * (t - 0.8)^2)))/1.35 
```

### Constant variance

We first look at the case of constant variance.

```{r poisson-constant-variance}
sigma.ini = sqrt(var1)
sigma.t = sigma.ini/mean(sigma.ini) * sd(mu.t)/rsnr^2
set.seed(327)
X.s = matrix(rnorm(10 * n, mu.t, sigma.t), nrow = 10, byrow = TRUE)
mu.est = apply(X.s, 1, smash.gaus)
mu.est.tivar.ash = apply(X.s, 1, ti.thresh, method = "smash")
mu.est.tivar.mad = apply(X.s, 1, ti.thresh, method = "rmad")
mu.est.ti = matrix(0, 10, n)
mu.est.ti.ebayes = matrix(0, 10, n)
for (i in 1:10) {
    sig.est = sig.est.func(X.s[i, ], n)
    mu.est.ti[i, ] = waveti.u(X.s[i, ], noise.level = sig.est)
    mu.est.ti.ebayes[i, ] = waveti.ebayes(X.s[i, ], noise.level = sig.est)
}
```

Assess accuracy of the results.

```{r calc-mise-poisson-1}
cat("SMASH:",mise(t(mu.est), mu.t, 10),"\n")
cat("TI thresholding with variance estimated from smash:",
    mise(t(mu.est.tivar.ash), mu.t, 10),"\n")
cat("TI thresholding with variance estimated from running MAD:",
    mise(t(mu.est.tivar.mad), mu.t, 10),"\n")
cat("TI thresholding with constant variance (estimated):",
	mise(mu.est.ti, mu.t, 10),"\n")
cat("EBayes with constant variance (estimated):",
	mise(mu.est.ti.ebayes, mu.t, 10),"\n")
```

Assess accuracy over a grid.

```{r calc-mise-grid-poisson-1}
xgrid = 1:n/n
mu.est.inter = matrix(0, 10, n)
mu.est.tivar.ash.inter = matrix(0, 10, n)
mu.est.tivar.mad.inter = matrix(0, 10, n)
mu.est.ti.inter = matrix(0, 10, n)
mu.est.ti.ebayes.inter = matrix(0, 10, n)
for (i in 1:10) {
    mu.est.inter[i, ] = approx(t, mu.est[, i], xgrid)$y
    mu.est.tivar.ash.inter[i, ] = approx(t, mu.est.tivar.ash[, i], xgrid)$y
    mu.est.tivar.mad.inter[i, ] = approx(t, mu.est.tivar.mad[, i], xgrid)$y
    mu.est.ti.inter[i, ] = approx(t, mu.est.ti[i, ], xgrid)$y
    mu.est.ti.ebayes.inter[i, ] = approx(t, mu.est.ti.ebayes[i, ], xgrid)$y
}
mu.t.inter = (spike.f(xgrid) + 1)/5
mise(mu.est.inter, mu.t.inter, 10)
mise(mu.est.tivar.ash.inter, mu.t.inter, 10)
mise(mu.est.tivar.mad.inter, mu.t.inter, 10)
mise(mu.est.ti.inter, mu.t.inter, 10)
mise(mu.est.ti.ebayes.inter, mu.t.inter, 10)
```

Plot the estimated mean functions against the ground-truth function
(in black).

```{r plots-6, fig.width=5, fig.height=4}
par(mfrow = c(1, 1))
plot(t, mu.t, xlab = "", ylab = "", type = "l")
lines(t, mu.est[, 1], col = 2)
lines(t, mu.est.tivar.mad[, 1], col = 3)
lines(t, mu.est.ti[1, ], col = 4)
lines(t, mu.est.ti.ebayes[1, ], col = 6)
legend("topright",legend = c("smash", "ti_rmad", "ti_homo", "ebayes_homo"),
       fill = c(2, 3, 4, 6))
```

### Non-constant variance

We then look at the case of non-constant variance.

```{r poisson-nonconstant-var}
sigma.ini = sqrt(var2)
sigma.t = sigma.ini/mean(sigma.ini) * sd(mu.t)/rsnr^2
set.seed(327)
X.s = matrix(rnorm(10 * n, mu.t, sigma.t), nrow = 10, byrow = TRUE)
mu.est = apply(X.s, 1, smash.gaus)
mu.est.tivar.ash = apply(X.s, 1, ti.thresh, method = "smash")
mu.est.tivar.mad = apply(X.s, 1, ti.thresh, method = "rmad")
mu.est.ti = matrix(0, 10, n)
mu.est.ti.ebayes = matrix(0, 10, n)
for (i in 1:10) {
    sig.est = sig.est.func(X.s[i, ], n)
    mu.est.ti[i, ] = waveti.u(X.s[i, ], noise.level = sig.est)
    mu.est.ti.ebayes[i, ] = waveti.ebayes(X.s[i, ], noise.level = sig.est)
}
```

Assess accuracy of the results.

```{r calc-mise-poisson-2}
cat("SMASH:",mise(t(mu.est), mu.t, 10),"\n")
cat("TI thresholding with variance estimated from smash:",
    mise(t(mu.est.tivar.ash), mu.t, 10),"\n")
cat("TI thresholding with variance estimated from running MAD:",
    mise(t(mu.est.tivar.mad), mu.t, 10),"\n")
cat("TI thresholding with constant variance (estimated):",
	mise(mu.est.ti, mu.t, 10),"\n")
cat("EBayes with constant variance (estimated):",
	mise(mu.est.ti.ebayes, mu.t, 10),"\n")
```

Assess accuracy over a grid.

```{r calc-mise-grid-poisson-2}
xgrid = 1:n/n
mu.est.inter = matrix(0, 10, n)
mu.est.tivar.ash.inter = matrix(0, 10, n)
mu.est.tivar.mad.inter = matrix(0, 10, n)
mu.est.ti.inter = matrix(0, 10, n)
mu.est.ti.ebayes.inter = matrix(0, 10, n)
for (i in 1:10) {
    mu.est.inter[i, ] = approx(t, mu.est[, i], xgrid)$y
    mu.est.tivar.ash.inter[i, ] = approx(t, mu.est.tivar.ash[, i], xgrid)$y
    mu.est.tivar.mad.inter[i, ] = approx(t, mu.est.tivar.mad[, i], xgrid)$y
    mu.est.ti.inter[i, ] = approx(t, mu.est.ti[i, ], xgrid)$y
    mu.est.ti.ebayes.inter[i, ] = approx(t, mu.est.ti.ebayes[i, ], xgrid)$y
}
mu.t.inter = (spike.f(xgrid) + 1)/5
mise(mu.est.inter, mu.t.inter, 10)
mise(mu.est.tivar.ash.inter, mu.t.inter, 10)
mise(mu.est.tivar.mad.inter, mu.t.inter, 10)
mise(mu.est.ti.inter, mu.t.inter, 10)
mise(mu.est.ti.ebayes.inter, mu.t.inter, 10)
```

```{r plots-7, fig.width=5, fig.height=4}
par(mfrow = c(1, 1))
plot(t, mu.t, xlab = "", ylab = "", type = "l")
lines(t, mu.est[, 1], col = 2)
lines(t, mu.est.tivar.mad[, 1], col = 3)
lines(t, mu.est.ti[1, ], col = 4)
lines(t, mu.est.ti.ebayes[1, ], col = 6)
legend("topright", legend = c("smash", "ti_rmad", "ti_homo", "ebayes_homo"),
       fill = c(2, 3, 4, 6)) 
```

## Simulation example from Fan & Yao (1998)

Now we look at some of the simulations in a couple of papers, as well
as the datasets used in them. The first one is from
[Fan & Yao (1998)](https://doi.org/10.1093/biomet/85.3.645).
The simulation is as described in Example 2. To deal with the fact
that the sample size is not a sample size of 2, we first reflect the
right portion of the data about the right endpoint so that it is a
power of 2. To make the data periodic, we then reflect the new data
about the right endpoint again. Our results are shown below.

```{r fan-yao}
mu.mad = 0
var.mad = 0

cat("Running 400 simulations.\n")
for (i in 1:400) {
    cat(i,"")
    x = sort(runif(200, -2, 2))
    err = rnorm(200)
    sigma.t = 0.4 * exp(-2 * x^2) + 0.2
    mu.t = x + 2 * exp(-16 * x^2)
    y = mu.t + sigma.t * err
    
    xgrid = seq(-1.8, 1.8, length.out = 101)
    mu.t.inter = xgrid + 2 * exp(-16 * xgrid^2)
    var.t.inter = (0.4 * exp(-2 * xgrid^2) + 0.2)^2
    
    y.exp = c(y, y[200:145])
    y.data = c(y.exp, y.exp[256:1])
    
    mu.est = smash.gaus(y.data)
    mu.est = mu.est[1:200]
    var.est = smash.gaus(y.data, v.est = TRUE)
    var.est = var.est[1:200]
    
    mu.est.inter = approx(x, mu.est, xgrid, "linear")$y
    var.est.inter = approx(x, var.est, xgrid, "linear")$y
    
    mu.mad[i] = 1/101 * sum(abs(mu.est.inter - mu.t.inter))
    var.mad[i] = 1/101 * sum(abs(var.est.inter - var.t.inter))
}
cat("\n")	
```

*Add text here.*

```{r}
boxplot(var.mad)

plot(x, mu.t, xlab = "", ylab = "", type = "l", main = "mean function")
lines(x, mu.est[1:200], col = 2)

plot(x, sigma.t^2, xlab = "", ylab = "", type = "l", main = "var function")
lines(x, var.est[1:200], col = 2) 
```

## Treasury Bill example from Fan & Yao (1998)

We also apply SMASH to the dataset in Example 1 of the Fan & Yao
(1998) paper. To take into account replicates, we take the median of
the response at the same point The results are shown below.

```{r, eval=FALSE}
data(treas)

library(tseries)

y = ar(data, FALSE, 5)$res
y = y[!is.na(y)]
x = sort(data[5:(length(data) - 1)])
y = y[order(data[5:(length(data) - 1)])]

plot(x, y)

## average replicate data

x.mod = unique(x)
y.mod = 0
for (i in 1:length(x.mod)) {
    y.mod[i] = median(y[x == x.mod[i]])
}

y.exp = c(y.mod, y.mod[length(y.mod):(2 * length(y.mod) - 2^10 + 1)])
y.final = c(y.exp, y.exp[length(y.exp):1])

y.est = smash.gaus(y.final)
y.est.var = smash.gaus(y.final, v.est = TRUE) 

```

```{r fig.width=12, fig.height=12, eval=FALSE}
par(mfrow = c(2, 2))
plot(data, xlab = "year", ylab = "interest rate", type = "l")

plot(x, y)
lines(x.mod, y.est[1:(length(y.mod))], xlab = "X", ylab = "Y", col = 2)

plot(x.mod, y.est[1:(length(y.mod))], xlab = "X", ylab = "Y", type = "l", ylim = c(-0.3, 0.3))

plot(x.mod, sqrt(y.est.var[1:(length(y.mod))]), xlab = "X", ylab = "volatility", type = "l")
````

We now look at the paper by Delouille et al. (2004). Specifically, we
compare with the heavisine function used in section 6, as the actual
function is readily available. The results for $n=200$ for both the
homoskedastic and teh heteroskedastic cases are given.

```{r, eval=FALSE}
n = 200
mse.hsm.c = 0
mse.hsm.n = 0
for (i in 1:500) {
    xt = sort(rnorm(n, 0.5, 0.2))
    xt = (xt - min(xt))/(max(xt) - min(xt))
    
    mu.hsm = 4 * sin(4 * pi * xt) - 2 * sign(xt - 0.3) - 2 * sign(0.72 - xt)
    
    mu.t = mu.hsm
    var1 = rep(1, n)
    var2 = (xt <= 0.5) + 2 * (xt > 0.5)
    sigma.ini = sqrt(var1)
    rsnr = sqrt(4)
    sigma.t = sigma.ini/mean(sigma.ini) * sd(mu.t)/rsnr^2
    
    y = rnorm(n, mu.t, sigma.t)
    y.exp = c(y, y[200:145])
    y.final = c(y.exp, y.exp[256:1])
    
    mu.est <- smash.gaus(y.final)
    mu.est = mu.est[1:n]
    mse.hsm.c[i] = mse(mu.est, mu.t)
    
    sigma.ini = sqrt(var2)
    rsnr = sqrt(4)
    sigma.t = sigma.ini/mean(sigma.ini) * sd(mu.t)/rsnr^2
    
    y = rnorm(n, mu.t, sigma.t)
    y.exp = c(y, y[200:145])
    y.final = c(y.exp, y.exp[256:1])
    
    mu.est <- smash.gaus(y.final)
    mu.est = mu.est[1:n]
    mse.hsm.n[i] = mse(mu.est, mu.t)
    
}

sqrt(quantile(mse.hsm.c, seq(0.25, 0.75, 0.25)))
sqrt(quantile(mse.hsm.n, seq(0.25, 0.75, 0.25))) 
```

We also look at the motorcycle in the same paper. We plot the data
points and fit, as well as the estimated variances.

```{r, eval=FALSE}
library(MASS)

x.ini = sort(mcycle$times)
y.ini = mcycle$accel[order(mcycle$times)]

x = unique(x.ini)
y = 0
for (i in 1:length(x)) {
    y[i] = median(y.ini[x.ini == x[i]])
}

y.exp = c(y, y[length(y):(2 * length(y) - 128 + 1)])
y.final = c(y.exp, y.exp[length(y.exp):1])

y.est = smash.gaus(y.final)
y.est = y.est[1:length(y)]
y.var.est = smash.gaus(y.final, v.est = TRUE)
y.var.est = y.var.est[1:length(y)] 
```
